Polynomial Time and Dependent Types

### 2023-10-24 paper/paper.tex
[ paper ] fixes in related work

\paragraph{Explicit Resource Accounting with Dependent Types} In
contrast to the implicit systems, previous works have constructed
systems that give explicit resource bounds via typing. Examples
include \citet{HoffmannDW17}'s Resource Allocated ML (RAML) and
\citet{RajaniG0021}, both of which are based on ideas of type-based
amortised complexity analysis arising from \cite{hofmann99lfpl}'s
ideas, via the work of \cite{HofmannJ03}. More details are to be found
in the survey paper of \citet{HoffmannJ22}. A approach is to track
costs at the value level instead of the types. \citet{Danielsson08}
describes a system that uses a ``tick'' effect to count steps of
computation, which can be reasoned about via dependent
types. \citet{NiuSGH22} take this idea further by employing a
modality-based phase separation to ensure that tick counting never
interferes with the functional business of
programs. \citet{McCarthyFNFF16} is another tick effect based system
in Coq. All of these techniques rely on the programmer correctly
annotating the program with tick effects to count the resource usage
they are interested in, in contrast to the intrinsic approach we have
taken here.

\paragraph{Linear and Substructural Dependent Types} We chose QTT as
the particular combination of linear and dependent types for our
systems. Other systems include systems such as those by
\cite{CervesatoP02,KrishnaswamiPB15,Vakar14} which all use a strict
separation between linear and non-linear variables. This strict
separation would mean that we could not as easily move programs from
the linear fragment into the types, as in
\autoref{sec:programming-polytime}. Systems that are more like QTT in
that they do not have a strict separation of variables include those
of \cite{MoonEO21,ChoudhuryEEW21,Abel18}. These systems differ from
QTT in that they do not include a complete copy of unrestricted type
theory as QTT does in its $\sigma = 0$ fragment, because they all
track usage in types as well as terms, so it is not clear how to use
them for unrestricted reasoning as we do with QTT. \citet{FuKS22}
present a system that is closer to QTT but does not include a universe
type.

>>>>

\paragraph{Explicit Resource Accounting with Dependent Types} In
contrast to the implicit systems, previous works have constructed
systems that give explicit resource bounds via typing. Examples
include \citet{HoffmannDW17}'s Resource Allocated ML (RAML) and
\citet{RajaniG0021}, both of which are based on ideas of type-based
amortised complexity analysis arising from \cite{hofmann99lfpl}'s
ideas, via the work of \citet{HofmannJ03}. More details are to be
found in the survey paper of \citet{HoffmannJ22}. Another approach is
to track costs at the value level instead of the
types. \citet{Danielsson08} describes a system that uses a ``tick''
effect to count steps of computation, which can be reasoned about via
dependent types. \citet{NiuSGH22} take this idea further by employing
a modality-based phase separation to ensure that tick counting never
interferes with the functional business of
programs. \citet{McCarthyFNFF16} is another tick effect based system
in Coq. All of these tick-counting techniques rely on the programmer
correctly annotating the program with tick effects to count the
resource usage they are interested in, in contrast to the intrinsic
approach we have taken here.

\paragraph{Linear and Substructural Dependent Types} We chose QTT as
the particular combination of linear and dependent types for our
systems. Other systems include systems such as those by
\citet{CervesatoP02}, \citet{KrishnaswamiPB15}, and \citet{Vakar14}
which all use a strict separation between linear and non-linear
variables. This strict separation would mean that we could not as
easily move programs from the linear fragment into the types, as in
\autoref{sec:programming-polytime}. Systems that are more like QTT in
that they do not have a strict separation of variables include those
of \citet{MoonEO21}, \citet{ChoudhuryEEW21}, and \citet{Abel18}. These
systems differ from QTT in that they do not include a complete copy of
unrestricted type theory as QTT does in its $\sigma = 0$ fragment,
because they all track usage in types as well as terms, so it is not
clear how to use them for unrestricted reasoning as we do with
QTT. \citet{FuKS22} present a system that is closer to QTT but does
not include a universe type, which we used in \autoref{sec:classes} to
be able to characterise complexity classes as predicates decidable in
restricted complexity.
